What we {\sf eval} in the shadows

### 2021-04-01 analysis/environments.Rmd
Envir

```{r}
envir_env <- E_envir %>% filter(envir_type == "environment") 
envir_unusual <- E_envir %>% filter(envir_type %in% c("list", "integer", "double", "NULL", "pairlist"))
envir_incorrect <- E_envir %>% anti_join(envir_env) %>% anti_join(envir_unusual)

>>>>

```{r}
envir_env <- E_envir %>% filter(envir_type == "environment") 
# S4 class: from pim package, uses an environment under the hood
envir_unusual <- E_envir %>% filter(envir_type %in% c("list", "integer", "double", "NULL", "pairlist", "S4"))
envir_incorrect <- E_envir %>% anti_join(envir_env) %>% anti_join(envir_unusual)

### 2021-04-01 paper/main.tex
Envir

\item \underline{$min(e)=\c{X}$}: Variables are the most common form. They
  subsume values, thus \c{X} includes arithmetic expressions in which a variable
  occurs. \packageNbSymbolVarSitePercent of variable expression are symbols, \ie
  just an \c{x}. As a variable may refer to a promise, it may trigger
  computation.
  %% The
  %% \c{envir} argument is explicitly passed in \packageMinimizedpercentenvira\% of
  %% the call sites. The call sites without an explicit \c{envir} argument all use
  %% a specific \eval variant, \c{eval.parent}. Moreover, the \c{envir} argument is
  %% both explicit and different from \c{parent.frame()} in
  %% \packageMinimizedpercentenvirnonparenta\% of the sites, which suggests that
  %% when given a variable, \eval is often used to get value in a non-local,
  %% non-parent environment.
\item \underline{$min(e)=\c{\$}$}: A name lookup, \c{x\$f}, can be used to
  access variable in an environment or a column in a data frame, this also
  include vector   indexing operators, \c{[]} and \c{[[]]}. A name lookup
  subsumes variables.
%% It amounts for \packageNbSlotAccess sites, \ie
%% \packageSlotAccessPercent of the sites. The non-default, non-parent environment
%% percentage is here \packageMinimizedpercentenvirnonparentg.

\item \underline{$min(e)=$~\c{<-}}: Assignments account for
  \packageAssignSitesPercent of \eval sites, they are \c{<-}, {\tt <<-} and
  \c{assign}, and they subsume named lookups.
%% \packageNonDefaultEnvirAssignSitesPercent of the sites with assignments. Indeed,
%% most of those assignments occur in the default environment, \ie in the caller of
%% \eval. This hints that side effects are not widespread in \eval.\footnote{Side
%% effects can also happen in a function call in the \eval expression, and that
%% function call will hide assignment operators from us. We more precisely
%% instrument side-effects to environment in Section~REF.}

\item \underline{$min(e)=\c{F()}$}: A plain function call which may have some
  values as argument but does not include variables. It should be noted that
  function name require variable lookups -- we distinguish those as they usually
  do not trigger computation.

\item \underline{$min(e)=\c{F(X)}$}: Function calls whose arguments may include
  variable references.

\item \underline{$min(e)=\c{F(F(X))}$}: Nested function calls with variables.

\item \underline{$min(e)=\c{FUN}$}: Function definition accounts for
  \packageFunctionDefinitionSitesPercent of \eval sites. In addition the
  functions inside other expressions account for
  \packageGeneralizedFunctionDefinitionSitesPercent of sites, and give an
  estimate of the use of higher-order functions.

\item \underline{$min(e)=\c{\{\,\}}$}: Arbitrary-length statement blocks. These
  are larger expressions with, on average, \packageMinimizedoperationsjRnd
  instructions.

\item \underline{$min(e)=\c{model.frame}$}: Call to the \c{model.frame} function
  used to create statistical models.

@@

@@

\begin{table}[htbp]{ \centering
\begin{tabular}{c|c|c|c}\hline
\multicolumn{2}{c|}{Core} & \multicolumn{2}{|c}{Packages}  \\\hline
Environment Class & Eval Calls \% & Environment Class & Eval Calls\\\hline
\CoreEnvClassA & \CoreProportionA & \PackagesEnvClassA &  \PackagesProportionA\\
\CoreEnvClassB & \CoreProportionB & \PackagesEnvClassB &  \PackagesProportionB\\
\CoreEnvClassC & \CoreProportionC & \PackagesEnvClassC &  \PackagesProportionC\\
\CoreEnvClassD & \CoreProportionD & \PackagesEnvClassD &  \PackagesProportionD\\
\CoreEnvClassE & \CoreProportionE & \PackagesEnvClassE &  \PackagesProportionE\\\hline
\end{tabular}} \label{tab:environments}
\caption{Environments in terms of \eval calls}
\end{table}

@@

\paragraph{\c{enclos} argument.}

>>>>

\item \underline{$min(e)=\c{X}$}: Variables are the most common normal form.
  Variables subsume values, so they include arithmetic expressions in which a
  variable occurs. \packageNbSymbolVarSitePercent of variable expression are
  symbols, \ie just an \c{x}. A variable may refer to a promise and thus trigger
  computation the first time it is accessed.
  %% The
  %% \c{envir} argument is explicitly passed in \packageMinimizedpercentenvira\% of
  %% the call sites. The call sites without an explicit \c{envir} argument all use
  %% a specific \eval variant, \c{eval.parent}. Moreover, the \c{envir} argument is
  %% both explicit and different from \c{parent.frame()} in
  %% \packageMinimizedpercentenvirnonparenta\% of the sites, which suggests that
  %% when given a variable, \eval is often used to get value in a non-local,
  %% non-parent environment.
\item \underline{$min(e)=\c{\$}$}: A name lookup, \c{x\$f}, is used
  can be used
  to look up a name in an environment or a column in a data frame. This category
  also groups indexing operators, \c{[]} and \c{[[]]}. The presence of these
  operators subsumes constants and variables.
%% It amounts for \packageNbSlotAccess sites, \ie
%% \packageSlotAccessPercent of the sites. The non-default, non-parent environment
%% percentage is here \packageMinimizedpercentenvirnonparentg.

\item \underline{$min(e)=$~\c{<-}}: Assignments amount to
  \packageAssignSitesPercent of the sites, they include \c{<-}, \c{<<-} and
  \c{assign}, and subsume named lookups.
%% \packageNonDefaultEnvirAssignSitesPercent of the sites with assignments. Indeed,
%% most of those assignments occur in the default environment, \ie in the caller of
%% \eval. This hints that side effects are not widespread in \eval.\footnote{Side
%% effects can also happen in a function call in the \eval expression, and that
%% function call will hide assignment operators from us. We more precisely
%% instrument side-effects to environment in Section~REF.}

\item \underline{$min(e)=\c{F()}$}: Function calls that

\item \underline{$min(e)=\c{F(X)}$}: Function calls that

\item \underline{$min(e)=\c{F(F(X))}$}: Function calls that

\item \underline{$min(e)=\c{FUN}$}: Function calls that

\item \underline{$min(e)=\c{\{\,\}}$}: Function calls that

\item \underline{$min(e)=\c{model.frame}$}: Function calls that

@@

\paragraph{\c{model.frame}}
This is a function used to create statistical models. We explain it in more
details in the section about the usage of \eval (see Section~REF).

\paragraph{Function definition.}
\c{FUN} represents a function definition at the top level of the expression, and
happens in \packageFunctionDefinitionSites sites, \ie
\packageFunctionDefinitionSitesPercent of the sites. It entails very few steps
of the interpreter, as the function is returned by \eval, and the body of the
new function is just stored for later execution. If we count also function
definition inside more complex expressions, it amounts to
\packageGeneralizedFunctionDefinitionSites sites, \ie
\packageGeneralizedFunctionDefinitionSitesPercent and suggests the widespread
use of higher-order functions in R, such as \c{lapply}, the R version of
\c{map}.

\paragraph{Large expressions.}  Compositions of functions, without variables
(\c{F(F())} minimized expression) is the second most frequent minimized
expression. \c{\{\}} is a sequence of statements. It is usually large, with
in average \packageMinimizedoperationsjRnd operations in it.

@@

% This table is old and incorrect
% Working on updating it
\begin{table}[htbp]{ \centering
\begin{tabular}{c|c|c|c}\hline
\multicolumn{2}{c|}{Core} & \multicolumn{2}{|c}{Packages}  \\\hline
Environment Class & Eval Calls \% & Environment Class & Eval Calls\\\hline
\CoreEnvClassA & \CoreProportionA & \PackagesEnvClassA &  \PackagesProportionA\\
\CoreEnvClassB & \CoreProportionB & \PackagesEnvClassB &  \PackagesProportionB\\
\CoreEnvClassC & \CoreProportionC & \PackagesEnvClassC &  \PackagesProportionC\\
\CoreEnvClassD & \CoreProportionD & \PackagesEnvClassD &  \PackagesProportionD\\
\CoreEnvClassE & \CoreProportionE & \PackagesEnvClassE &  \PackagesProportionE\\\hline
\end{tabular}} \label{tab:environments}
\caption{Environments in terms of \eval calls}
\end{table}

@@

\paragraph{\c{enclos} argument.} \ \\
